---
title: "createFlatFilesRunscript"
author: "Sam Graf"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
whole_doc_startTime <- Sys.time()

library(tidyverse)
library(dataRetrieval)
library(readxl)
library(sf)
library(DT) #for tables in QAQC runscript

neededFunctions <- c("getDailyand15MinUSGSData.R", "getStatusFunction.R")

for (i in neededFunctions) {
  source(paste0("./functions/",i))
}

antennaMetadata <- read_excel("data/antennaMetadata.xlsx")

detections <- read_csv("data/detections_20251216.csv", 
    col_types = cols(dec_tag = col_character(), 
                     antenna = col_character(),
                     detected = col_character()))
detections <- detections %>%
  mutate(detected = ymd_hms(detected)) #%>%
  # add_row(`dec_tag` = "989.101030620161", 
  #         detected = as_datetime("2025-11-11 00:02:53"), 
  #         antenna = "02")

Unc_Tag_Releases <- read_csv("data/Unc Tag Release File Feb26.csv", 
                             col_types = cols(`Dec Tag #` = col_character())) # %>%
  # add_row(`Dec Tag #` = "989.00103062016133") %>%
  # add_row(`Dec Tag #` = "989.001030620161") %>%
  # add_row(`Dec Tag #` = "989.101030620161")

antennasSFAll <- st_as_sf(antennaMetadata, coords = c("long", "lat"), crs = 4326) 
antennasSF <- antennasSFAll %>%
  distinct(geometry, .keep_all = TRUE)
```



```{r USGS, include=FALSE, echo=FALSE}
studyStartDate <- min(date(detections$detected))

USGSFlows <- getDailyand15MinUSGSData("09147025", startDate = min(date(detections$detected), na.rm = TRUE), waterTemp = FALSE)

###Data Wrangling
#gets rid of redundant columns and rows
Unc_Tag_Releases1 <- Unc_Tag_Releases %>%
  select(`Release Date` = `Date 1st Enc`, Species, `TL 1st Enc. (mm)` = `TL 1st Enc`, `Dec Tag #`) %>%
  filter(!is.na(`Dec Tag #`))

detectionsAttributesFlows <- detections %>%
  left_join(antennasSFAll, by = c("antenna" = "antennaNumber"
  )) %>%
  left_join(Unc_Tag_Releases1, by = c("dec_tag" = "Dec Tag #")) %>%
  mutate(DetectionDate = date(detected)) %>%
  left_join(USGSFlows$USGSDataDaily, by = c("DetectionDate" = "Date"))
  #st_as_sf()

dailyStatus <- getStatusFunction(detectionsAttributesFlows, studyStartDate = studyStartDate)

combinedDetectionAndStatusData <- list(
  "detectionsAttributesFlows" = detectionsAttributesFlows,
  "dailyStatus" = dailyStatus
)


```

```{r qaqc messaging, echo=FALSE, include=FALSE}
###if something different than 16 is displayed, printhte tables
#####QAQC
#qaqc to make sure tag numbers brought in correctly; should all be length 16 (3 digits and decimal +12 digits proceeding)
UncTagDataCharacterTagcounts <- Unc_Tag_Releases %>%
  distinct(`Dec Tag #`) %>%
  count(`String Length of Released Tag Entries` = str_length(`Dec Tag #`), name = "Count") 
UncTagDataCharacterTagcountsFlagged <- UncTagDataCharacterTagcounts %>%
  filter(`String Length of Released Tag Entries` != 16)

detectionsTagCounts <- detections %>%
  distinct(dec_tag) %>%
  count(`String Length of all Detected Tag Entries` = str_length(dec_tag), name = "Count") 
detectionsTagCountsFlagged <- detectionsTagCounts %>%
  filter(`String Length of all Detected Tag Entries` != 16)
#files currently displayed in app
# morethan1dec_tag <- detectionsAttributesFlows %>%
#   distinct(dec_tag, newTag) %>%
#   count(`Release File tag entry` = newTag, name = "Number of dec_tag Entries") %>%
#   filter(`Number of dec_tag Entries` > 1)

moreThan1ReleaseEntry <- Unc_Tag_Releases1 %>%
  count(`Dec Tag #`, name = "Number of Tag Entries") %>%
  filter(`Number of Tag Entries` > 1)

detectionsWithoutReleaseData <- detectionsAttributesFlows %>%
  filter(is.na(`Release Date`))

##flag data and display if anything is out of the ordinary
if(nrow(UncTagDataCharacterTagcountsFlagged) > 0){
  problemReleaseTagsLengthMessage <- paste("Tags in the Release Data are not the correct format. Length of all Dec Tag entries should be 16: 3 digits before decimal, 12 subsequent digits.")
  problemReleaseDatatable <-  datatable(UncTagDataCharacterTagcounts, 
            rownames = FALSE,
            caption = "String Lengths of Tags Release Data (should be 16)"
            )
} else{
  problemReleaseTagsLengthMessage <- ""
  problemReleaseDatatable <- NULL
}

if(nrow(detectionsTagCountsFlagged) > 0){
  problemDetectionTagsLengthMessage <- paste("Tags in the Detection Data are not the correct format. Length of all Dec Tag entries shoult be 16: 3 digits before decimal, 12 subsequent digits.")
  problemDetectionDatatable <- datatable(detectionsTagCounts, 
            rownames = FALSE,
            caption = "String Lengths of Tags in Detection Data (should be 16)"
            )
} else {
  problemDetectionTagsLengthMessage <- ""
  problemDetectionDatatable <- NULL
}
# more than one release data tag entry
if(nrow(moreThan1ReleaseEntry) > 0){
  moreThan1ReleaseEntryMessage <- paste("There are tags with duplicate entries in the release file.")
  moreThan1ReleaseEntryDatatable <- datatable(moreThan1ReleaseEntry, 
            rownames = FALSE,
            caption = "Tags in release file that have more than 1 entry in the release file."
            )
} else {
  moreThan1ReleaseEntryMessage <- ""
  moreThan1ReleaseEntryDatatable <- NULL
}

# more than one dec tag entry in combined data
if(nrow(detectionsWithoutReleaseData) > 0){
  detectionsWithoutReleaseDataMessage <- paste("There are detected tags without release data from the release file.")
  detectionsWithoutReleaseDataDatatable <- datatable(detectionsWithoutReleaseData, 
            rownames = FALSE,
            caption = "Detected tags without release data."
            )
} else {
  detectionsWithoutReleaseDataMessage <- ""
  detectionsWithoutReleaseDataDatatable <- NULL
}

qaqcData <- list(#"morethan1dec_tag" = morethan1dec_tag, 
                 "moreThan1ReleaseEntry" = moreThan1ReleaseEntry, 
                 "detectionsWithoutReleaseData" = detectionsWithoutReleaseData)
```

<span style="color:red; font-size: 30px;" >`r problemReleaseTagsLengthMessage`</span>
`r problemReleaseDatatable`

<span style="color:red; font-size: 30px;" >`r problemDetectionTagsLengthMessage`</span>
`r problemDetectionDatatable`

<span style="color:red; font-size: 30px;" >`r moreThan1ReleaseEntryMessage`</span>
`r moreThan1ReleaseEntryDatatable`

<span style="color:red; font-size: 30px;" >`r detectionsWithoutReleaseDataMessage`</span>
`r detectionsWithoutReleaseDataDatatable`
```{r savingFiles, echo=FALSE, include=TRUE}
allFlatFiles <- list(
  "combinedDetectionAndStatusData" = combinedDetectionAndStatusData, 
  "USGSFlows" = USGSFlows,
  "antennasSF" = antennasSF,
  "qaqcData" = qaqcData
)

for(filename in names(allFlatFiles)){
  saveRDS(allFlatFiles[[filename]], paste0("data/flatFilesforApp/", filename, ".rds"))
}

whole_doc_endTime <- Sys.time()

```

## `r paste("The whole document took", round(difftime(whole_doc_endTime, whole_doc_startTime, units = "mins"),2), "minutes to run.  Flat Files are saved to the flat files directory and the app is ready to run with the most updated data.")`